---
title: "Project_1"
author: "Aaron Bench"
date: "March 4, 2019"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

## Data Ingestion 


```{r, echo=TRUE}
setwd("C:/Users/aaron/OneDrive/Documents/Data_Science")
#Ingest data

df <- read.csv("activity.csv")

df$date<- as.Date(df$date)

#create subset, removing missing values

dfs <- subset(df, steps >= 0)  

```

##What is mean total number of steps taken per day?

### 1. Total Number of steps taken per day

```{r echo=TRUE}
p1 <- ggplot(data=dfs, aes(x=date, y=steps)) + 
  geom_line (color="#00AFBB", size = 1)
p1
```
This graph shows total staps taken per day (excluding data points with NA values)

### 2. Histogram

```{r, echo=TRUE}
p2 <- ggplot(data=dfs, aes(x=steps)) + 
  geom_histogram ( color="#00AFBB", binwidth = 10) + geom_vline(aes(xintercept = mean(steps)), color="red", linetype="dashed", size = 1) +geom_vline(aes(xintercept=median(steps)), color="blue", linetype="dashed", size= 1) + geom_density()
p2

```

This guy doesn't walk much.  Blue line is the median of 0.  Red line is an average of 37.38

###3. Mean and Median

```{r, echo=TRUE}
mean(dfs$steps)
median(dfs$steps)

```
##What is the average daily activity pattern?

###1. Time Series Plot

```{r echo=TRUE}
###Average steps  per day

avg <- tapply(dfs$steps, dfs$interval, mean, na.rm=TRUE, simplify=T)
dfas <- data.frame(interval=as.integer(names(avg)), avg=avg)


int_plot <- ggplot(data=dfas, aes(x=interval, y=avg)) + 
  geom_line (color="#00AFBB", size = 1)
int_plot



```




Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


###5 minute interval, max average across all days


```{r echo=TRUE}
ms<- max(dfas$avg)
dfas[dfas$avg == ms,]

```
####Interval 835 has the max average steps of 206

##Inputting Missing Values 

### 1.Number of missing values 

```{r echo=TRUE}
sum(is.na(df$steps))
```

####2304 NA's in steps data

###2-3.Strategy to fill in missing data and create a new data set 
```{r echo=TRUE}

for (i in 1:length(df$steps)) {  #If NA use mean(steps) else use populated data
  if(is.na(df$steps[i])) {
    df$stepsb[i] <- mean(df$steps, na.rm=T)
  }
  else {
    df$stepsb[i] <- df$steps[i]
  }
}


```

###4.Hostogram of daily steps with populated missing data

```{r echo=TRUE}

dfsum <- aggregate(df$stepsb, by = list(df$date), FUN = sum) #subset of data with days summmed 

names(dfsum)[names(dfsum)=="x"] <- "Total"

names(dfsum)[names(dfsum)=="Group.1"] <- "Date"

mean(df$stepsb)

median(df$stepsb)
     
#VS

mean(dfs$steps)
median(dfs$steps)

hist2 <- ggplot(data = dfsum, aes(Total)) + 
    geom_histogram(binwidth = 1500, color = "blue") +
    xlab("Total Number of Steps Taken per Day") +
    ylab("Count") +
    ggtitle("Total Steps per Day - mean used to replace NULL values")
print(hist2)


```
#### When I use average to fill in NULL/NA values, the mean and median do not change from the subset where NA/NULL values are removed.

##Differences in Activity Pattern by week days

###1. Factor Variable for Weekday
```{r echo=TRUE}
df$wkday <- ifelse(weekdays(df$date) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), "W-day", "W-end")

head(df)

```
###2. Panel plot
```{r echo=TRUE}

new_df <- (df %>% group_by(interval, wkday) %>% summarise(Mean = mean(stepsb)))

g <- ggplot(new_df, mapping = aes(x = interval, y = Mean)) + geom_line(color="blue") +
    facet_grid(wkday ~.) + xlab("Interval") + ylab("Mean of Steps") +
    ggtitle("Weekday VS. Weekend") 

g

g2 <- ggplot(new_df, mapping = aes(x = interval, y = Mean)) + geom_boxplot(outlier.colour="black", outlier.shape=16,
               outlier.size=2, notch=FALSE, aes(fill = wkday)) 
g2

```
####There are differences.


